### 2021-06-15 JSS/using_drtmle_jss.Rmd
finish first pass

We note that this method of computing confidence intervals can also be used to
obtain a transformed confidence interval for a particular marginal mean. For
example, recall that $Y$ is binary in our running example. Thus, we may wish to
enforce that our confidence interval be computed on a scale ensuring that the
estimated confidence limits fall within the unit interval $[0, 1]$. To that end,
we can utilize an interval of the form:
\begin{equation*}
  \expit \biggl[ \log\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
  \pm z_{1-\alpha/2} \sigma^{\text{logit}}_n \biggr] \ .
\end{equation*}
This confidence interval may be implemented as follows.

@@

In our examples so far, we have only considered binary treatments. However,
`drtmle` is equipped to handle treatments with an arbitrary number of discrete
values. Suppose that $A$ assumes values in a set $\mathcal{A}$, the PS
regression estimation is modified to ensure that
\begin{equation*}
  \sum\limits_{a \in \mathcal{A}} g_n(a \mid w) = 1 \ \mbox{for all $w$} \ .
\end{equation*}
If this condition is true, we say that the estimates of the PS are compatible.
Many regression methodologies that have been developed work well with binary
outcomes, but have not been extended to multi-level outcomes. To ensure that
users of `drtmle` have access to the large number of binary regression
techniques when estimating the PS, we have taken an alternative approach to
estimation of the propensity for each level of the treatment. Specifically,
rather than fitting a single multinomial-type regression, we fit a series of
binomial regressions. As a concrete example, consider the case of no missing
data and where $\mathcal{A} = \{0,1,2\}$. We can ensure compatible estimates of
the PS by generating an estimate $g_n(0 \mid w)$ of $Pr_0(A = 0 \mid W = w)$
and $\tilde{g}_n(1 \mid w)$ of $Pr_0(A = 1 \mid A > 0, W = w)$. The latter
estimate may be generated by regression $I(A = 1)$ on $W$ in observations with
$A > 0$. Because the outcome is binary, this regression may be estimated using
any technique suitable for binary outcomes. By simple rules of conditional
probability, we know that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
Pr_0(A = 0 \mid W)\}$ and thus an estimate of $g_0(1 \mid w)$ can be computed
as $g_n(1 \mid w) = \tilde{g}_n(1 \mid w) \{1 - g_n(0 \mid w)\}$. Finally, we
let $g_n(2 \mid w) = 1 - g_n(0 \mid w) - g_n(1 \mid w)$, which ensures
compatibility of the estimates for each level of covariates. This approach
generalizes to an arbitrary number of treatment levels. Below, we provide an
illustration using simulated data with a treatment that has three levels. The
true parameter values in this simulation are $\psi_0(0)=$ `r round(EY0,2)`,
$\psi_0(1)=$ `r round(EY1, 2)`, and $\psi_0(2)=$ `r round(EY2,2)`.

@@

Alternatively, we can specify a `contrast` to estimate confidence intervals
about the average treatment effect of $A=1$ vs. $A=0$. Calls to `wald_test` can
also be made.

@@

Similarly, we can compute confidence intervals comparing the average treatment
effect of $A = 2$ vs. $A = 0$.

@@

Similarly, we can compute confidence intervals for the risk ratio comparing the
marginal mean for $A=2$ to $A=1$.

@@

When considering adaptive estimation techniques, one runs the risk of
overfitting the initial regressions. While the cross validation used by super
learning attempts to guard against overfitting, there are no guarantees in
finite-samples. To definitively avoid such overfitting, one can use
cross-validated estimates of the regression parameters. Theoretically, the use
of cross-validated regression weakens the assumptions necessary to achieve
asymptotic normality [@zheng:2010:working]. This is true of the regular TMLE and
AIPTW, as well as of the TMLE estimator with doubly-robust inference. The
`drtmle` function implements cross-validated estimation of the regression
parameters through the `cvFolds` option, as shown in the code below, where we
continue with the multi-level treatment example.

@@

Cross-validation may significantly increase the computation time necessary for
running `drtmle`. Parallelized regression fitting is available by setting the
option `use_future = TRUE`. Since, internally, all computations are carried out
using futures (with the API provided by the "future" R package), setting this
option amounts only to parallelizing with futures [@futurePackage] via
`future.apply::lapply`. Selecting this option results in parallelization of the
estimation of the OR, PS, and reduced-dimension regressions.

If no adjustments are made prior to invoking the `drtmle` function, futures are
computed sequentially. It is up to the user to specify appropriate system-specific 
future back-ends that may better suit their problem (e.g., using the `future.batchtools` 
package [@futurebatchtoolsPackage]). Below we
demonstrate registration using an ad-hoc cluster on the local system. The choice
of back-end is flexible, and submission of the job to a wide variety
of schedulers (e.g., SLURM, TORQUE, etc.) is also permitted. Interested readers
are invited to consult the documentation of the `future.batchtools` package for
further information.

```{r, cache = TRUE, message = FALSE, eval = FALSE}
## Commented out to avoid build errors
# library(future.batchtools)
# library(parallel)
# cl <- makeCluster(2, type = "SOCK")
# plan(cluster, workers = cl)
# clusterEvalQ(cl, library("SuperLearner"))
# pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
#                      SL_g = c("SL.glm", "SL.glm.interaction","SL.earth"),
#                      SL_Q = c("SL.glm", "SL.glm.interaction","SL.earth"),
#                      SL_gr = c("SL.glm", "SL.earth", "SL.mean"),
#                      SL_Qr = c("SL.glm", "SL.earth", "SL.mean"),
#                      stratify = FALSE, a_0 = c(0,1,2),
#                      cvFolds = 2, use_future = TRUE)
```

@@

Doubly-robust estimators of marginal means are appealing in their robustness and
efficiency properties. However, researchers may prefer the IPTW estimator for
its intuitive derivation and implementation. Heretofore, inference for IPTW
estimators precluded the use of adaptive estimators of the PS. However, \citet{vdl:2014:ijb} proposed modified IPTW estimators that allow for adaptive PS
estimation. Similarly as above, this estimator is based on an additional
regression parameter, $\tilde{Q}_{r,0n}(a,w) := E_0\{Y \mid A = a, g_n(W) =
g_n(w)\}$, that is the regression of the outcome $Y$ on the estimated propensity
amongst observations with $A=a$. The author showed that if $g_n^*(a \mid w)$, an
estimator of the propensity for treatment $A=a$, satisfied that
$$ \frac{1}{n} \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W =
W_i)} \ \{ I(A_i = a) - g_n(a \mid W = W_i)\} = 0 \ ,$$
then the IPTW estimator based on $g_n^*$ has an asymptotic normal distribution
with an estimable variance. In fact, \citet{vdl:2014:ijb} proposed a TMLE
algorithm that mapped an initial PS estimate into an estimate that satisfies
this key equation and variance estimators that can be used to generate
Wald-style confidence intervals and hypothesis tests.

An estimator with equivalent asymptotic properties based on a PS estimate,
$g_n(a \mid w)$, can be computed as
$$
  \psi_{n,IPTW}(a) = \psi_{n,IPTW}(a) - \frac{1}{n} \sum\limits_{i=1}^n
  \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W = W_i)} \ \{ I(A_i = a) - g_n(a
  \mid W = W_i)\} \ .
$$
Both this estimator and the TMLE-based IPTW estimator above are implemented in
`drtmle`.

@@

The IPTW estimators based on adaptive estimation of the PS are implemented in
the `adaptive_iptw` function. The `adaptive_iptw` function shares many options
with `drtmle`. However, as the IPTW estimator does not rely on an estimate of
the OR, these options are omitted. We continue with the multilevel treatment
example.

@@

The `"adaptive_iptw"` class contains identical methods for computing confidence
intervals and Wald tests as the `"drtmle"` class and we refer readers back to
confidence intervals and hypothesis tests section for reminders on the various
options for these tests. By default, these methods are implemented for the
`iptw_tmle` estimator of \citet{vdl:2014:ijb}. We expect that this estimator
will have improved finite-sample behavior relative to `iptw_os` (the
alternative estimator described in the previous subsection), though further
study is needed.

As with `drtmle`, the `adaptive_iptw` function allows missing data in `A` and
`Y`. Similarly, `adaptive_iptw` also allows for parallelized, cross-validated
estimation of the regression parameters via `cvFolds` and `parallel` options.
As with the estimators implemented in `drtmle`, cross-validation allows for
valid inference under weaker assumptions.

@@

We have recently proposed 
estimators of the average treatment effect to be used in settings where this 
effect is weakly identifiable, as evidenced by small estimated propensity scores. 
In these settings doubly robust estimators may exhibit non-robust behavior
in small samples. The proposed solution was to target an alternative quantity 
in the propensity estimation step. If we are interested in estimating $\psi_0(a)$
for $a = 0, 1$, we can estimate $Pr_0(A = a \mid \bar{Q}_0(1, W), \bar{Q}_0(0, W))$,
which describes the conditional probability of receiving treatment $A = a$ given
the value of the two outcome regressions. For this reason, we call this quantity
an *outcome-adaptive propensity score*. We have shown that all of the usual 
asymptotics for TMLE carry through when an estimator of this quantity is 
substituted for the true propensity score. The resultant TMLE is *super-efficient*
meaning that it will generally have greater asymptotic efficiency than a standard
TMLE. Simulations indicate that this approach also delivers more stable behavior
in settings where estimated propensities become very small. However, the estimator 
is not doubly robust -- the asymptotics rely on the outcome regression being
consistently estimated at $n^{-1/4}$ rate. In this sense, this estimator trades off
robustness in the name of achieving greater stability and efficiency. See our
paper for further discussion.

@@

If the outcome-adaptive PS is desired, the user may specify this via
the `adapt_g` option. In this case, the additional targeting steps needed for doubly
robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed above,
this estimator is *not* doubly robust. Note that when the `adapt_g = TRUE` the data
used for the PS estimation will be a `data.frame` with column names
`QaW` where `a` takes all the values specified in `a_0`. For example if `a_0 = c(0,1)`,
then the PS estimation data frame will have two columns called `Q0W`
and `Q1W`. This may be relevant for users writing their own `SuperLearner` wrappers
or those who make use of the `glm_g` setting.

@@

The `drtmle` package was designed to provide a user-friendly implementation of
the TMLE algorithm that facilitates doubly-robust inference about marginal
means and average treatment effects. While the theory underlying doubly-robust
inference is complex, the implementation of the methods only requires users to
provide the data and specify how to estimate regressions. While estimation of
the OR and PS is familiar to those familiar with causal inference-related
estimation, the reduced-dimension regressions may appear strange. Indeed, it is
difficult to provide intuition for these parameters and more difficult to
determine what scientific background knowledge might guide users in how to
estimate these regression parameters. Therefore, we recommended estimating
these quantities adaptively, e.g., with the super learner. The super learner
library should contain several relatively smooth estimators such as `SL.mean`,
`SL.glm`, and `SL.gam`. In particular, including `SL.mean` (an unadjusted
generalized linear model) may be important due to the fact that, when the OR
and PS are consistently estimated, the reduced-dimension regressions are
identically equal to zero. Thus, this unadjusted regression estimator may do a
good job estimating the reduced-dimension regressions in situations where the
OR and PS are estimated well. In addition to relatively smooth estimators, we
recommend additionally including several adaptive estimators such as `SL.npreg`
or `SL.earth`.

Planned extensions of the `drtmle` package include extensions of longitudinal
settings involving treatments at multiple time points and inclusion of the
collaborative targeted maximum likelihood estimator (C-TMLE) of
\citet{vdl:2014:ijb} that also facilitates collaborative doubly-robust
inference.

>>>>

We note that this method of computing confidence intervals can also be used to
obtain a transformed confidence interval for a particular marginal mean. For
example, recall that $Y$ is binary in our running example. Thus, we may wish to
enforce that our confidence interval be computed on a scale ensuring that the
estimated confidence limits fall within the unit interval $[0, 1]$. To that end,
we can utilize an interval of the form:
\begin{equation*}
  \text{expit} \biggl[ \log\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
  \pm z_{1-\alpha/2} \sigma^{\text{logit}}_n \biggr] \ .
\end{equation*}
This confidence interval may be implemented as follows.

@@

In our examples so far, we have only considered binary treatments. However,
\pkg{drtmle} is fully equipped to handle treatments with an arbitrary number of
discrete values as well. Suppose that $A$ assumes values in a set $\mathcal{A}$
of limited cardinality, the PS regression estimation is modified to ensure that
\begin{equation*}
  \sum\limits_{a \in \mathcal{A}} g_n(a \mid w) = 1 \ \mbox{for all $w$} \ .
\end{equation*}
If this condition is true, we say that the estimates of the PS are compatible.
Many regression methodologies that have been developed work well with binary
outcomes, but have not been extended to multi-level outcomes. To ensure that
users of \pkg{drtmle} have access to the large number of binary regression
techniques when estimating the PS, we have taken an alternative approach to
estimation of the propensity for each level of the treatment. Specifically,
rather than fitting a single multinomial-type regression, we fit a series of
binomial regressions. As a concrete example, consider the case of no missing
data and where $\mathcal{A} = \{0,1,2\}$. We can ensure compatible estimates of
the PS by generating an estimate $g_n(0 \mid w)$ of $Pr_0(A = 0 \mid W = w)$
and $\tilde{g}_n(1 \mid w)$ of $Pr_0(A = 1 \mid A > 0, W = w)$. The latter
estimate may be generated by regressing $I(A = 1)$ on $W$ in observations with
$A > 0$. Since the outcome is binary, this regression may be estimated using
any technique suitable for binary outcomes. By simple rules of conditional
probability, it holds that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
Pr_0(A = 0 \mid W)\}$ and thus an estimate of $g_0(1 \mid w)$ can be computed
as $g_n(1 \mid w) = \tilde{g}_n(1 \mid w) \{1 - g_n(0 \mid w)\}$. Finally, we
let $g_n(2 \mid w) = 1 - g_n(0 \mid w) - g_n(1 \mid w)$, which ensures
compatibility of the estimates for each level of the covariates. This approach
generalizes to an arbitrary number of treatment levels. Below, we provide an
illustration using simulated data with a treatment that has three levels. The
true parameter values in this simulation are $\psi_0(0)=$ `r round(EY0,2)`,
$\psi_0(1)=$ `r round(EY1, 2)`, and $\psi_0(2)=$ `r round(EY2,2)`.

@@

Alternatively, we can specify a `contrast` to estimate confidence intervals
about the average treatment effect of $A=1$ versus $A=0$. Calls to `wald_test()`
can also be made.

@@

Similarly, we can compute confidence intervals comparing the average treatment
effect of $A = 2$ versus $A = 0$.

@@

Just as straightforwardly, we can compute confidence intervals for the risk
ratio comparing the marginal mean for $A=2$ to $A=1$.

@@

When considering adaptive estimation techniques, one runs the risk of
overfitting the initial nuisance regressions. While the layer of
cross-validation built into super learning attempts to guard against
overfitting, there are no guarantees in finite samples. To definitively avoid
such overfitting, one can use cross-validated estimates of the nuisance
regressions. Theoretically, the use of cross-validated regression weakens the
assumptions necessary to achieve asymptotic normality [@zheng:2010:working].
This is true of the regular TML and AIPTW estimators, as well as of the TML
estimator with doubly-robust inference. The `drtmle()` function implements
cross-validated estimation of the regression parameters through the `cvFolds`
option, as shown in the code below, where we continue on with the multi-level
treatment example.

@@

Cross-validation may significantly increase the computation time necessary for
running `drtmle()`. Parallelized nuisance regression fitting is available by
setting the option `use_future = TRUE`. Since, internally, all computations are
carried out using futures (with the API provided by the \pkg{future} R package),
setting this option amounts only to parallelizing with futures [@futurePackage]
via `future.apply::lapply()`; when `use_future = FALSE`, base `lapply()` is
simply used instead.  Together, setting these options results in parallelization
of the estimation of the OR, PS, and reduced-dimension regressions.

If no adjustments are made prior to invoking the `drtmle()` function, futures
are computed sequentially, as per the defaults of the \pkg{future} package. It
is left to the user to specify appropriate system-specific future backends that
may better suit their problem (e.g., using the \pkg{future.batchtools} package
[@futurebatchtoolsPackage] or even simply setting a parallelized
`future::plan()`). Below, we demonstrate registration using an ad-hoc cluster on
the local system. The choice of backend is flexible, and submission of the job
to a wide variety of schedulers (e.g., SLURM, TORQUE) is also permitted.
Interested readers are invited to consult the documentation of the
\pkg{future.batchtools} package for further information.

```{r, message = FALSE, eval = FALSE}
library(future.batchtools)
library(parallel)
cl <- makeCluster(2, type = "SOCK")
plan(cluster, workers = cl)
clusterEvalQ(cl, library("SuperLearner"))
pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_g = c("SL.glm", "SL.glm.interaction","SL.earth"),
                     SL_Q = c("SL.glm", "SL.glm.interaction","SL.earth"),
                     SL_gr = c("SL.glm", "SL.earth", "SL.mean"),
                     SL_Qr = c("SL.glm", "SL.earth", "SL.mean"),
                     stratify = FALSE, a_0 = c(0,1,2),
                     cvFolds = 2, use_future = TRUE)
```

@@

Doubly-robust estimators of marginal means are appealing in their robustness and
efficiency properties. However, researchers may prefer the IPTW estimator for
its intuitive derivation and implementation. Heretofore, inference for IPTW
estimators precluded the use of adaptive estimators of the PS. However,
\citet{vdl:2014:ijb} proposed modified IPTW estimators that allow for adaptive
PS estimation. Similarly as above, this estimator is based on an additional
regression parameter, $\tilde{Q}_{r,0n}(a,w) \coloneqq E_0\{Y \mid A = a, g_n(W)
= g_n(w)\}$, that is, the regression of the outcome $Y$ on the estimated
propensity score amongst observations with $A=a$. The author showed that if
$g_n^{\star}(a \mid w)$, an estimator of the propensity for treatment $A=a$,
satisfied that
\begin{equation*}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W =
    W_i)} \ \{ I(A_i = a) - g_n(a \mid W = W_i)\} = 0 \ ,
\end{equation*}
then the IPTW estimator based on $g_n^{\star}$ would have an asymptotic normal
distribution with estimable variance. In fact, \citet{vdl:2014:ijb} proposed a
TMLE algorithm that mapped an initial PS estimate into an estimate that
satisfies this key equation and variance estimators that can be used to generate
Wald-style confidence intervals and hypothesis tests.

An estimator with equivalent asymptotic properties based on a PS estimate,
$g_n(a \mid w)$, can be computed as
\begin{equation*}
  \psi_{n,IPTW}(a) = \psi_{n,IPTW}(a) - \frac{1}{n} \sum\limits_{i=1}^n
  \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W = W_i)} \ \{ I(A_i = a) - g_n(a
  \mid W = W_i)\} \ .
\end{equation*}
Both this estimator and the TMLE-based IPTW estimator above are implemented in
\pkg{drtmle}.

@@

The IPTW estimators based on adaptive estimation of the PS are implemented in
the `adaptive_iptw()` function. The `adaptive_iptw()` function shares many
options with `drtmle()`; however, as the IPTW estimator does not rely on an
estimate of the OR, options related to the OR are omitted. We demonstrate below,
while continuing with the multi-level treatment example.

@@

The `"adaptive_iptw"` S3 class contains identical methods for computing
confidence intervals and Wald-style tests as the `"drtmle"` class, thus readers
are referred back to the section on confidence intervals and hypothesis tests
for reminders on the various options for these procedures. By default, these
methods are implemented for the `iptw_tmle()` estimator of \citet{vdl:2014:ijb}.
We conjecture that this estimator will have improved finite-sample behavior
relative to `iptw_os()` (the alternative estimator described above), though
further study in numerical experiments is warranted.

As with `drtmle()`, the `adaptive_iptw()` function allows missing data in `A`
and `Y`. Similarly, `adaptive_iptw()` also allows for parallelized,
cross-validated estimation of the regression parameters via the `cvFolds` and
`parallel` options.  As with the estimators implemented by `drtmle()`,
cross-validation allows for valid inference under weaker assumptions.

@@

We have recently proposed estimators of the average treatment effect to be used
in settings where this effect is weakly identifiable, as evidenced by estimated
propensity scores of small magnitude.  In these settings, doubly robust
estimators may exhibit non-robust behavior in small samples. The proposed
solution was to target an alternative quantity in the propensity score
estimation step. If we are interested in estimating $\psi_0(a)$ for $a = 0, 1$,
we can estimate $Pr_0(A = a \mid \bar{Q}_0(1, W), \bar{Q}_0(0, W))$, which
describes the conditional probability of receiving treatment $A = a$ given the
value of the two outcome regressions. For this reason, we call this quantity an
*outcome-adaptive propensity score*. We have shown that all of the usual
asymptotics for TMLE carry through when an estimator of this quantity is
substituted for the true propensity score. The resultant TML estimator is
*super-efficient*, meaning that it will generally have greater asymptotic
efficiency than a standard TML estimator. Simulation experiments indicate that
this approach also delivers more stable behavior in settings where estimated
propensity scores become very small. However, the estimator is not doubly robust
-- the asymptotics rely on the outcome regression being consistently estimated
at $n^{-1/4}$ rate. In this sense, this estimator trades off robustness against
achieving greater stability and efficiency. See \citet{ju2020robust} for further
discussion.

@@

If the outcome-adaptive PS is desired, the user may specify this via the
`adapt_g` option. In this case, the additional targeting steps needed for doubly
robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed
above, this estimator is *not* doubly robust. Note that when `adapt_g = TRUE`,
the data used for PS estimation will be a `data.frame` with column names `QaW`
where `a` takes all the values specified in `a_0`. For example if `a_0 =
c(0,1)`, then the PS estimation data frame will have two columns called `Q0W`
and `Q1W`. This may be relevant for users writing their own \pkg{SuperLearner}
wrapper functions or those who make use of the `glm_g` setting.

@@

The \pkg{drtmle} package was designed to provide a user-friendly implementation
of the TMLE algorithm that facilitates doubly-robust inference about marginal
means and average treatment effects. While the theory underlying doubly-robust
inference is complex, the implementation of the methods only requires users to
provide the data and specify how to estimate nuisance regressions. Though
estimation of the OR and PS is familiar to those with backgrounds in causal
inference-related estimation, the reduced-dimension regressions may appear
strange. Indeed, it is difficult to provide intuition for these parameters and
more difficult still to determine what scientific background knowledge might
guide users in how to estimate these regression parameters. Therefore, we
recommended estimating these quantities adaptively, e.g., with the super
learner. The super learner library should contain several relatively smooth
estimators such as `SL.mean`, `SL.glm`, and `SL.gam`. In particular, including
`SL.mean` (an unadjusted generalized linear model) may be important due to the
fact that, when the OR and PS are consistently estimated, the reduced-dimension
regressions are identically equal to zero. Thus, this unadjusted regression
estimator may do a good job estimating the reduced-dimension regressions in
situations where the OR and PS are estimated well. In addition to relatively
smooth estimators, we recommend additionally including several adaptive
estimators such as `SL.npreg` or `SL.earth`.

Planned extensions of the \pkg{drtmle} package include extensions to
longitudinal settings involving treatments at multiple time points and inclusion
of the collaborative targeted maximum likelihood estimator (C-TMLE) of
\citet{vdl:2014:ijb} that also facilitates collaborative doubly-robust
inference.
